---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---


##Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
str(activity)
```

We don't want the date to be a factor variable

```{r}
activity$date <- as.Date(activity$date)
str(activity)
summary(activity)
```

###What is mean total number of steps taken?

```{r}
hist(tapply(activity$steps[!is.na(activity$steps)], activity$date[!is.na(activity$steps)], sum), 10, col="red", xlab = "Steps", main = "Histogram of steps taken each day")
```

Lets calculate the mean using the same tapply

```{r}
mean(tapply(activity$steps[!is.na(activity$steps)], activity$date[!is.na(activity$steps)], sum), na.rm = T)
```

Now the median

```{r}
median(tapply(activity$steps[!is.na(activity$steps)], activity$date[!is.na(activity$steps)], sum), na.rm = T)
```

##What is the average daily activity pattern

```{r}
dap <- tapply(activity$steps[!is.na(activity$steps)], activity$interval[!is.na(activity$steps)], mean)
plot(as.integer(names(dap)), dap, type = "l", col = "red", main = "Average steps taken by interval", xlab= "Interval", ylab= "Average steps taken")
```

Now lets see the 5-minute interval that contains the maximum number of steps

```{r}
names(dap[which(dap==max(dap))])
```

##Imputing missing values

Lets see how many missing values we have in our dataset (we could also look that it in the summary)

```{r}
sum(is.na(activity$steps))
```

To fill the missing values, I particularly used the interval mean we calculated previously

```{r}
activity2 <- data.frame(date = activity$date, interval = activity$interval)
for (i in 1:length(activity$steps)) {
        if (is.na(activity$steps[i])) {
                activity2$steps[i] <- dap[which(activity$interval[i] == as.integer(names(dap)))]
        }
        else {
                activity2$steps[i] <- activity$steps[i]        
        }
}
```

Lets summarize our new dataset with no NA's values

```{r}
str(activity2)
summary(activity2)
```

Histogram:

```{r}
hist(tapply(activity2$steps[!is.na(activity2$steps)], activity2$date[!is.na(activity2$steps)], sum), 10, col="red", xlab = "Steps", main = "Histogram of steps taken each day")
```

Mean:

```{r}
mean(tapply(activity2$steps, activity2$date, sum))
```

Median:

```{r}
median(tapply(activity2$steps, activity2$date[!is.na(activity2$steps)], sum))
```

Given the strategy used the stimates remain almost the same (using averages wont affect averages but totals). 

##Are there differences in activity patterns between weekdays and weekends?

Lets create a new factor variable in the dataset with two levels, “weekday” and “weekend”

```{r}
for (i in 1:length(activity2$date)) {
        if (weekdays(activity2$date[i])%in%c("Saturady","Sunday")) {
                activity2$day[i] <- "weekend"
        }
        else {
                activity2$day[i] <- "weekday"        
        }
}
```

Now we create a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days

```{r}
wda <- tapply(activity2$steps, list(activity2$interval, activity2$day), mean)
wda2 <- data.frame(steps = append(wda[,1],wda[,2]), interval=as.numeric(c(names(wda[,1]), names(wda[,1]))))
for (i in 1:288){wda2$day[i] <- "weekday"}
for (i in 1:288){wda2$day[288+i] <- "weekend"}
library(lattice)
xyplot(steps ~ interval | day, wda2, main="Time series panel plot of the 5-minute intervals", xlab="Interval",ylab="Average steps", type = "l",layout = c(1,2))
```

